Different forms of uncertainty differently affect the evolution of social learning

### 2022-08-01 ms.tex
Updates from Overleaf

Empirically, the number of learning opportunities can be manipulated in the lab,
but will also correspond with an individual's age the real world. In multi-round
studies of information use in novel tasks, US participants' use of social
information declined precipitously across rounds ~\cite{McElreath2005}, suggesting
they were more likely to use social information when they were most uncertain
about the task early on. In a more naturalistic context, Aplin et al. (2017)
\nocite{Aplin2017} found that younger great tits more readily used social
information compared with older individuals, possibly because they had accrued
less information via individual learning, and also possibly because younger
individuals have the most to gain (given their expected remaining lifespan) by
switching to superior behavioral options. 
% \ps{I got rid of the Leris and Reader reference because it didn't seem relevant
% to age-based uncertainty. They just showed that young guppies exposed to
% reliable models were more likely to use social information later. I also added
% some human studies, which seems important.} 
Cross-cultural studies on humans have shown the importance of childhood as a phase
of heavy social learning ~\cite{Reyes2016}. Young children are more likely to
acquire their beliefs and simple skills from their parents than are older children
or adults, which is at least partly due to the differential knowledge accrual
between young children and the older adults to whom they direct the most trust and
attention \cite{kline2013teaching}.
%The age of individuals was found to have an effect on social learning in both
%\emph{parus major}~\cite{Aplin2017} and \emph{poecilia reticulata}
%(guppies)~\cite{Leris2016}. \ps{What effect? I'm not sure } \mt{Need better
%connection between lifespan and age, or better references to support use of
%lifespan}.
Individuals with shorter effective lifespans will end their lives with more
uncertainty about which behaviors are optimal, and thereby transmit some of that
uncertainty to anyone who learns from them. The effective lifespan for learning
varies across tasks, individuals, and species, yet most models assume one learning
opportunity per generation ~\cite{Feldman1996,Henrich1998, perreault2012bayesian}.
Some models do allow several cultural generations within 1 genetic generation
~\cite{Enquist2007}, but little formal theory explicitly examines the role of
uncertainty based on number of learning opportunities into our understanding of
the evolution of social learning.

@@

Many models of social learning assume little, or nothing, cognitive capabilities
of their agents. For example, a common modelling strategy compare the payoffs of
agents with different pure learning strategies (some will always engage in social
learning, and others in individual learning) while revealing nothing about the
cognition underlying individuals' decisions  ~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. Some more complicated learning strategies that make use of both
socially and individually learned information have been proposed
~\cite{Enquist2007, perreault2012bayesian}, but these are seldom designed to
reflect cognitive mechanisms for integrating information. Though such
simplifications are often useful, several authors have warned of the risks to
\emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016,
Kendal2018}, given such details may be theoretically meaningful. In order to
understand how social learning evolves as an adaptation to uncertainty, we need to
endow agents with a  biologically plausible cognition capable of integrating
information from various sources in a payoff-maximizing way. \cm{I don't
  understand this next sentence, but think it can be cut with no loss of meaning:
  In our model we do not impose any structure in randomness between behaviors, so
our cognitive model is only sensitive to total randomness.} To do so, we let
agents see the payoffs to behaviors and use the softmax function to convert these
observations to a probability distribution from which they strategically select
their own behavior. \cm{does this still make sense? and does it need a citation?}

>>>>

Empirically, the number of learning opportunities can be manipulated in the lab,
but will also correspond with an individual's age the real world. In multi-round
studies of information use in novel tasks, US participants' use of social
information declined precipitously across rounds ~\cite{McElreath2005}, suggesting
they were more likely to use social information when they were most uncertain
about the task early on. In a more naturalistic context, Aplin et al. (2017)
\nocite{Aplin2017} found that younger great tits more readily used social
information compared with older individuals, possibly because they had accrued
less information via individual learning, and also possibly because younger
individuals have the most to gain (given their expected remaining lifespan) by
switching to superior behavioral options. 
% \ps{I got rid of the Leris and Reader reference because it didn't seem relevant
% to age-based uncertainty. They just showed that young guppies exposed to
% reliable models were more likely to use social information later. I also added
% some human studies, which seems important.} 
Cross-cultural studies on humans have shown the importance of childhood as a phase
of heavy social learning ~\cite{Reyes2016}. Young children are more likely to
acquire their beliefs and simple skills from their parents than are older children
or adults, which is at least partly due to the differential knowledge accrual
between young children and the older adults to whom they direct the most trust and
attention \cite{kline2013teaching}.
%The age of individuals was found to have an effect on social learning in both
%\emph{parus major}~\cite{Aplin2017} and \emph{poecilia reticulata}
%(guppies)~\cite{Leris2016}. \ps{What effect? I'm not sure } \mt{Need better
%connection between lifespan and age, or better references to support use of
%lifespan}.
%Individuals with shorter effective lifespans will end their lives with more
%uncertainty about which behaviors are optimal, and thereby transmit some of that
%uncertainty to anyone who learns from them. 
The effective lifespan for learning
varies across tasks, individuals, and species, yet most models assume one learning
opportunity per generation ~\cite{Feldman1996,Henrich1998, perreault2012bayesian}.
Some models do allow several cultural generations within 1 genetic generation
~\cite{Enquist2007}, but little formal theory explicitly examines the role of
learning opportunities in social learning evolution models.

@@

Many models of social learning assume little, or nothing, cognitive capabilities
of their agents. For example, a common modelling strategy compare the payoffs of
agents with different pure learning strategies (some will always engage in social
learning, and others in individual learning) while revealing nothing about the
cognition underlying individuals' decisions  ~\cite{BoydRicherson1985, Rogers1988,
aoki2005}. Some more complicated learning strategies that make use of both
socially and individually learned information have been proposed
~\cite{Enquist2007, perreault2012bayesian}, but these are seldom designed to
reflect cognitive mechanisms for integrating information. Though such
simplifications are often useful, several authors have warned of the risks to
\emph{blackboxing}, or ignoring, cognitive mechanisms \cite[p. 658]{Heyes2016,
Kendal2018}, given such details may be theoretically meaningful. In order to
understand how social learning evolves as an adaptation to uncertainty, we need to
endow agents with a  biologically plausible cognition capable of integrating
information from various sources in a payoff-maximizing way. \cm{I don't
  understand this next sentence, but think it can be cut with no loss of meaning:
  In our model we do not impose any structure in randomness between behaviors, so
our cognitive model is only sensitive to total randomness.} To do so, we let
agents see the payoffs to behaviors and use the softmax function to convert these
observations to a probability distribution from which they strategically select
their own behavior. \cm{does this still make sense? and does it need a citation?} \mt{makes sense, maybe this par could be merged with above into one?}
